---
title: "segmut"
output: 
  github_document: default
  html_document: default
date: "`r Sys.Date()`"
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# Package description

<!-- badges: start -->

<!-- badges: end -->

The goal of segmut is to segment genome with mutations to segments with significantly different mutations density.

## Installation

You can install the development version of segmut from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
# Sys.unsetenv("GITHUB_PAT")
devtools::install_github("mishashe/segmut")
```

To load nessesary libraries:
```{r libraries, echo=TRUE, results='hide'}
# library(segmut)
library(DEoptim, quietly=T) # to find optimal break points using differential evolution algorithm
library(RColorBrewer, quietly=T) # to plot results
library(doParallel, quietly=T)
ncores <- 4 # set to desired number of cores for parallel computation
registerDoParallel(cores = ncores)
library(rlist)
library(KSgeneral)
```
To set parameters
```{r set parameters}
Kmin <- 50
```

## Example with known number of breaks ```n=2```

This is a basic example which shows you how to find positions of ```n=2``` breaks:

To generate example of genome of length ```L``` with vector of mutation locations ```muts```:
```{r genome example}
L <- 10000
muts <- sort(c(sample(1:3000,3000*0.1),sample(3001:8000,5000*0.18),sample(8001:10000,2000*0.12)))
```
To find optimal breaks locations given ```n=2``` number of breaks
```{r find optimal breaks}
res <- getBreaks(muts = muts, L = L, Kmin=Kmin, n=2)
```
To plot the results
```{r plot results,fig.width=10, fig.height=0.5}
breaks <- sort(c(0,res$optim$bestmem,L))
colors <- brewer.pal(name="Paired", n=length(breaks)-1)
par(mar=c(2,0,0,0))
plot(muts,rep(0,length(muts)),pch=".", cex = 1.5,ylim=c(-0.06,0.01),ylab="",xlab="", axes=F)
axis(side=1, at=c(0,3000,8000,L))
for (i in 1:(length(breaks)-1))
{
  lines(c(breaks[i],breaks[i+1]),c(-0.05,-0.05),col=colors[i], lwd=5)
}
```

## Example with unknown number of breaks

This is a basic example which shows you how to find optimal number of breaks and their locations:

To find optimal number of breaks
```{r find optimal number of breaks}
breaks <- getNumberBreaks(muts,L=max(muts)-min(muts)+1,Kmin=100,pThreshold=0.05)
```
To plot the results
```{r plot best results,fig.width=10, fig.height=0.5}
breaks0L <- sort(c(0,breaks,L))
colors <- brewer.pal(name="Paired", n=length(breaks0L)-1)
par(mar=c(2,0,0,0))
plot(muts,rep(0,length(muts)),pch=".", cex = 1.5,ylim=c(-0.06,0.01),ylab="",xlab="", axes=F)
axis(side=1, at=c(0,3000,8000,L))
for (i in 1:(length(breaks0L)-1))
{
  lines(c(breaks0L[i],breaks0L[i+1]),c(-0.05,-0.05),col=colors[i], lwd=5)
}
```
