---
title: "segmut"
output: 
  github_document: default
  html_document: default
date: "`r Sys.Date()`"
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# Package description

<!-- badges: start -->

<!-- badges: end -->

The goal of segmut is to segment genome with mutations to segments with significantly different mutations density.

## Installation

You can install the development version of segmut from [GitHub](https://github.com/) with:

``` r
# detach("package:segmut", unload=TRUE)
# install.packages("devtools")
# Sys.unsetenv("GITHUB_PAT")
# install.packages("/home/misha/Documents/Development/segmut/", repos = NULL, type = "source")
devtools::install_github("mishashe/segmut")
```

To load nessesary libraries:
```{r libraries, echo=TRUE, results='hide'}
library(segmut)
library(DEoptim, quietly=T) # to find optimal break points using differential evolution algorithm
library(RColorBrewer, quietly=T) # to plot results
library(stringr)
library(tidyverse)
library(scales)
library(ggExtra)
library(stringi)
```
To set parameters
```{r set parameters}
Kmin <- 100
```

To generate example of genome of length ```L``` with vector of mutation locations ```muts```:
```{r genome example}
L <- 4000
muts <- sort(c(sample(1:700,700*0.1),
               sample(701:2000,2300*0.2),
               sample(2001:3200,1200*0.15),
               sample(3201:4000,800*0.08)))
```

## Example with known number of breaks ```n=3```

This is a basic example which shows you how to find positions of ```n=3``` breaks:

To find optimal breaks locations given ```n=3``` number of breaks
```{r find optimal breaks}
res <- getBreaksChiSquare(muts = muts, L = L, Kmin=Kmin, n=3)
```
To plot the results
```{r plot results,fig.width=10, fig.height=0.5}
breaks0L <- sort(c(0,res$optim$bestmem,L))
colors <- brewer.pal(name="Paired", n=length(breaks0L)-1)
par(mar=c(2,0,0,0))
plot(muts,rep(0,length(muts)),pch=".", cex = 1.5,ylim=c(-0.06,0.01),ylab="",xlab="", axes=F)
axis(side=1, at=c(0,700,2000,3200,L))
for (i in 1:(length(breaks0L)-1))
{
  lines(c(breaks0L[i],breaks0L[i+1]),c(-0.05,-0.05),col=colors[i], lwd=5)
}
```

## Example with unknown number of breaks

This is a basic example which shows you how to find optimal number of breaks and their locations:

To find optimal number of breaks
```{r find optimal number of breaks}
breaks <- getNumberBreaksChiSquare(muts,L=L,Kmin=Kmin)
```
To plot the results
```{r plot best results,fig.width=10, fig.height=0.5, echo=TRUE, message=FALSE, warning=FALSE}
breaks0L <- sort(c(0,breaks,L))
colors <- brewer.pal(name="Paired", n=length(breaks0L)-1)
par(mar=c(2,0,0,0))
plot(muts,rep(0,length(muts)),pch=".", cex = 1.5,ylim=c(-0.06,0.01),ylab="",xlab="", axes=F)
axis(side=1, at=c(0,700,2000,3200,L))
for (i in 1:(length(breaks0L)-1))
{
  lines(c(breaks0L[i],breaks0L[i+1]),c(-0.05,-0.05),col=colors[i], lwd=5)
}
```

## Example with Escherichia coli (NZ_CP033020.1) vs. Salmonella enterica (NZ_AP026948) alignment
We align two bacterial genomes using nucmer and get all the alignment blocks and the mutations
```{r alignment,fig.width=10, fig.height=0.5, echo=TRUE, message=FALSE, warning=FALSE}
datadir <- "/home/misha/Documents/Development/segmut/data/"
file1 <- paste0(datadir,"Escherichia_coli_1.fasta")
file2 <- paste0(datadir,"Salmonella_enterica_1.fasta")
system(paste0("nucmer --mum --prefix=",datadir,"align ",file1," ",file2), intern = TRUE, wait=TRUE)
alignment <- system(paste0("show-aligns -w 500000 ", datadir,"align.delta NZ_CP033020.1 NZ_AP026948.1"), intern = TRUE, wait=TRUE)
alignment <- alignment[str_detect(alignment,"\\^")]
divergence <- sum(str_count(alignment,"\\^"))/sum(nchar(alignment))
divergences <- (str_count(alignment,"\\^"))/(nchar(alignment))
Ls <- nchar(alignment)
r <- stri_locate_all(pattern = '^', alignment, fixed = TRUE)
rs <- c()
for (i in order(-nchar(alignment))) rs <- c(rs,diff(r[[i]][,1]))
```
Plotting divergences and lengths of the alignment blocks (not segmented ones): 
```{r blocks,fig.width=10, fig.height=0.5, echo=TRUE, message=FALSE, warning=FALSE,fig.width=2, fig.height=2}
p <- ggplot(data=data.frame(Ls=Ls,divergences=divergences), aes(Ls, divergences)) + 
          geom_point(size=0.2) +
          theme_bw() +
          scale_x_continuous(trans='log10',breaks = 10^(-10:10), labels = trans_format("log10", math_format(10^.x))) +
          # scale_y_continuous(trans='log10',breaks = 10^(-10:10), labels = trans_format("log10", math_format(10^.x))) +
          ylab("average block divergence") + xlab("block length") 
  ggExtra::ggMarginal(p, type = "histogram")
```
In total there are `r length(Ls)` blocks with the total length of `r sum(Ls)` bp.

Plotting divergences and lengths of the alignment blocks (not segmented ones): 
```{r breaking to segments,fig.width=10, fig.height=0.5, echo=TRUE, message=FALSE, warning=FALSE,fig.width=2, fig.height=2}
Ks <- c()
taus <- c()
for (i in order(-nchar(alignment)))
{
  L <- nchar(alignment[[i]])
  rtemp <- r[[i]][,1]
  muts <- sort(rtemp + sample(0:0,length(rtemp),replace=TRUE))
  breaks <- getNumberBreaksChiSquare(muts,L=L,Kmin=1)
  breaks0L <- sort(c(0,breaks,L))
  for (j in 1:(length(breaks0L)-1))
  {
    Ks <- c(Ks, breaks0L[j+1]-breaks0L[j]+1)
    taus <- c(taus, sum(muts<breaks0L[j+1] & muts>breaks0L[j])/(breaks0L[j+1]-breaks0L[j]+1))
  }
  rs <- c(rs,diff(r))
}
```
Plotting divergences and lengths of the alignment blocks (not segmented ones): 
```{r plot blocks,fig.width=10, fig.height=0.5, echo=TRUE, message=FALSE, warning=FALSE,fig.width=2, fig.height=2}
p <- ggplot(data=data.frame(Ks=Ks,taus=taus), aes(Ks, taus)) + 
          geom_point(size=0.2) +
          theme_bw() +
          scale_x_continuous(trans='log10',breaks = 10^(-10:10), labels = trans_format("log10", math_format(10^.x))) +
          # scale_y_continuous(trans='log10',breaks = 10^(-10:10), labels = trans_format("log10", math_format(10^.x))) +
          ylab("segment divergence") + xlab("segment length") 
  ggExtra::ggMarginal(p, type = "histogram")
```

Calculating empirical and pseudotheoretical MLDs: 
```{r calc mld ,fig.width=10, fig.height=0.5, echo=TRUE, message=FALSE, warning=FALSE,fig.width=2, fig.height=2}
rB <- c(seq(0.5,35.5,3),10^seq(log10(35.5)+0.1,4.8,0.1))
rV <- 0.5*(rB[-length(rB)]+rB[-1])
pH <- hist(rs,breaks=rB,plot=FALSE)

mPT <- rV*0
for (ir in 1:length(rV)) 
{
  mPT[ir] <- sum((Ks>rV[ir])*(2*taus+(Ks-rV[ir])*taus^2)*exp(-rV[ir]*taus))
}
for (iK in 1:length(Ks)) 
{
  ir <- which(Ks[iK]<=rB[2:length(rB)] & Ks[iK]>=rB[1:(length(rB)-1)])
  mPT[ir] <- mPT[ir] + exp(-taus[iK]*Ks[iK])/diff(rB)[ir]
}
dat <- data.frame(rV=rV,mE=pH$counts/diff(rB),mPT=mPT)
```





Plotting empirical and pseudotheoretical MLDs: 
```{r mld,fig.width=10, fig.height=0.5, echo=TRUE, message=FALSE, warning=FALSE,fig.width=2, fig.height=2}
ggplot(data=dat, aes(rV, mE)) + 
          geom_point(size=0.2) +
          geom_line(aes(rV, mPT)) +
          theme_bw() +
          scale_x_continuous(limits = c(1e0, 1e3),trans='log10',breaks = 10^(-10:10), labels = trans_format("log10", math_format(10^.x))) +
          scale_y_continuous(limits = c(1e-3, 1e5),trans='log10',breaks = 10^(-10:10), labels = trans_format("log10", math_format(10^.x))) +
          xlab("r") + ylab("m(r)") 
```






