---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# segmut

<!-- badges: start -->

<!-- badges: end -->

The goal of segmut is to segment genome with mutations to segments with significantly different mutations density.

## Installation

You can install the development version of segmut from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
# Sys.unsetenv("GITHUB_PAT")
devtools::install_github("mishashe/segmut")
```

## Example

This is a basic example which shows you how to solve a common problem:

To load nessesary libraries:
```{r libraries, echo=FALSE}
library(segmut)
library(DEoptim)
library(RColorBrewer)
```
To generate example of genome of length L with vector of mutation locations muts:
```{r genome example}
L <- 10000
muts <- sort(c(sample(1:3000,400),sample(3001:8000,250),sample(8001:10000,400)))
```
To find optimal breaks given n=2 number of breaks
```{r find optimal breaks}
# res <- getBreaks(muts = muts, L = L, Kmin=0, n=2)
```
To plot the results
```{r plot results}
# breaks <- sort(c(0,res$optim$bestmem,L))
# colors <- brewer.pal(name="Paired", n=length(breaks)-1)
# plot(muts,rep(0,length(muts)),pch=".")
# for (i in 1:(length(breaks)-1))
# {
#   lines(c(breaks[i],breaks[i+1]),c(-0.05,-0.05),col=colors[i], lwd=5)
# }
```
